---
title: "Estimating Pi using a lightsaber battle"
author: "Ty Tuff"
date: '2019-03-14'
slug: estimating-pi-using-a-lightsaber-battle
tags:
- Pi
- Pi day
categories: pi day games
---

```{r, echo=FALSE}
library(retistruct)
library(png)
#setwd("/Users/Ty/Desktop/GitHub/Tuff_R_blog/Tuff_R_blog/content/post/pi_day/")
```


# Happy Pi Day everyone!! 

### Since the date today is the first three digits of Pi (3.14), I thought I would celebrate by deriving an approximation of Pi using a lightsaber battle. I will first build an army of saber-wielding critters in La Fountaine park and then make a few transects through those battles and count the number of contacts we make with critter-sabers on along the way. When we're done, we take a ratio of hits vs. misses to estimate Pi.

## We start in a futuristic world where urban wildlife have aquired light sabers and taken over Montreal. 

```{r, echo=FALSE, fig.cap="Future squirels battle for food scraps near homeless man's dome tent.", out.width = '100%'}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-10,10), ylim=c(-10,10))
#Get the plot information so the image will fill the plot box, and draw it
ima <- readPNG("animals-with-lightsabers-11.png")
lim <- par()
rasterImage(ima, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])
```


### To battle these savage beasts, we must learn to make our own light saber. 
```{r}

set.seed(1.1)
light_saber_length <- 20
number_of_animal_sabers_in_a_park <- 5000
distance_between_transects <- 40.1

point_1 <- c(-(light_saber_length/2),(light_saber_length/2))
point_2 <- c(0,0)
single_saber <-cbind(point_1, point_2)
rownames(single_saber)<- c("point_1","point_2")
colnames(single_saber)<- c("x","y")
single_saber
```

```{r, echo=FALSE, fig.cap="Represent our lightsaber as a line with no width. A single beam of light."}
par(mar=c(0,0,0,0))
# image from Wiki commons
ima <- readPNG("Lightsaber_blue_rotate.png")

plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-10,10), ylim=c(-10,10))
#Get the plot information so the image will fill the plot box, and draw it
lim <- par()
rasterImage(ima, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])
lines(single_saber, col=adjustcolor("#00bcff", 0.8), lwd=8)
```

## Then distribute the army of saber-wielding critters across the landscape.

### First we rotate around the origin to set the instantanious angle of each animal's individual saber.

```{r, echo=1:25}
  angle <- runif(1, min=1, max=360)*pi/180
  xy <- as.matrix(single_saber)
  
  # Rotation
  cos.angle <- cos(angle)
  sin.angle <- sin(angle)
  after_rotation <- xy %*% t(matrix(c(cos.angle, sin.angle, -sin.angle, 
        cos.angle), 2, 2))
  
after_rotation
```





```{r, echo=FALSE, fig.cap="We rotate the lightsaber around the origin to change its angle."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-10,10), ylim=c(-10,10))

lim <- par()
rasterImage(ima, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])

lines(single_saber, col=adjustcolor("#00bcff", 0.8), lwd=8)

abline(v=0)
abline(h=0)
lines(after_rotation, col=adjustcolor("#00bcff", 0.8), lwd=8)
```





Then we move (a.k.a. translate) the saber to a location on the landscape where 
the critter is located. 

```{r}
  # Translation
  translation.x <- runif(1, min=10, max=90)
  translation.y <- runif(1, min=10, max=90)
  translation <- matrix(
    c(translation.x,translation.x,translation.y,translation.y), 2, 2)
  new_coord <- after_rotation + translation
  colnames(new_coord) <- c("x","y")
  rownames(new_coord) <- c("Point_1","Point_2")
```


```{r, echo=FALSE}
new_coord <- after_rotation + matrix(c(50,50,50,50),2,2)
```


```{r, echo=FALSE, fig.cap="After the angle is set, translate the lightsaber to a random location on the battlefield."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-10,301), ylim=c(-10,301))

park <- readPNG("park.png")
lim2 <- par()
rasterImage(park, lim2$usr[1], lim2$usr[3], lim2$usr[2]-0.5, lim2$usr[4])

rasterImage(ima, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])

abline(v=0)
abline(h=0)

lines(single_saber, col=adjustcolor("#00bcff", 0.8), lwd=4)
lines(new_coord, col=adjustcolor("#00bcff", 1), lwd=4)
```


Create a function to rotate and distribute sabers to all the critters on the battlefield.

```{r}
critter_saber <- function(){
  single_saber <-cbind(c(-5,5),c(0,0))
  angle <- runif(1, min=1, max=360)*pi/180
  xy <- as.matrix(single_saber)
  
  # Rotation
  cos.angle <- cos(angle)
  sin.angle <- sin(angle)
  after_rotation <- xy %*% matrix(c(cos.angle, sin.angle, -sin.angle, 
        cos.angle), 2, 2, byrow=TRUE )
  
  # Translation
  translation.x <- runif(1, min=10, max=90)
  translation.y <- runif(1, min=10, max=90)
  translation <- matrix(
    c(translation.x,translation.x,translation.y,translation.y), 2, 2)
  new_coord <- after_rotation + translation
  colnames(new_coord) <- c("x","y")
  rownames(new_coord) <- c("Point_1","Point_2")
  
  return(new_coord)   
}

critter_saber()
```

Repeat it many times to create and army
```{r}
for(i in 1:number_of_animal_sabers_in_a_park){
critter_saber()
}
```

We assume that all the critters are battling vigously 
and indescriminintly with each other so they are randomly distributed across 
space and the angles of all the sabers are uniformly distributed. 

```{r, echo=FALSE, fig.cap="An army of lightsaber-wielding critters in the middle of La Fountaine park. This is the position of each critter's lightsaber at the moment we pass them on our transect."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-1,101), ylim=c(-1,101))

park <- readPNG("park.png")
lim <- par()
rasterImage(park, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])


for(i in 1:number_of_animal_sabers_in_a_park){
lines(critter_saber(), col=adjustcolor("#00bcff", 0.4), lwd=1)
}
```


```{r, echo=1}
surface_line <- function(line_height = 0){
  matrix(c(0,100,line_height,line_height), 2, 2, byrow=FALSE)
  }
surface_line(20)
```

```{r, echo=FALSE, fig.cap="My transect through the army of critters."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-1,101), ylim=c(-1,101))
park <- readPNG("park.png")
lim <- par()
rasterImage(park, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])


lines(surface_line(30), col=adjustcolor("white", 0.8), lwd=3)
```



```{r}

one_saber <- critter_saber()
one_surface_line <- surface_line(30)
line.line.intersection(one_saber[1,], one_saber[2,], one_surface_line[1,], 
                       one_surface_line[2,], interior.only = TRUE)
```


```{r, echo=FALSE, fig.cap="Build a mechanism to identify when a lightsaber has crossed my transect line."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-1,101), ylim=c(-1,101))
park <- readPNG("park.png")
lim <- par()
rasterImage(park, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])

a <- line.line.intersection(one_saber[1,], one_saber[2,], one_surface_line[1,], 
                            one_surface_line[2,], interior.only = TRUE)
lines(one_surface_line, col=adjustcolor("white", 0.8), lwd=3)
lines(one_saber, col=adjustcolor("#00bcff", 0.8), lwd=3)
points(a[1], a[2], cex=6, col="white")

one_saber_2 <- critter_saber()
one_surface_line_2 <- surface_line(30)
b <- line.line.intersection(one_saber_2[1,], one_saber_2[2,], 
                            one_surface_line_2[1,], one_surface_line_2[2,], 
                            interior.only = TRUE)
lines(one_saber_2, col=adjustcolor("#00bcff", 0.8), lwd=3)
points(b[1], b[2], cex=6, col="white")
```





```{r, echo=FALSE, fig.cap="Team prepares for transects", out.width = '100%'}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-20,20), ylim=c(-4,4))
#Get the plot information so the image will fill the plot box, and draw it
ima <- readPNG("wp2848056.png")
lim <- par()
rasterImage(ima, lim$usr[1], lim$usr[3], lim$usr[2], lim$usr[4])
```

Prepare a list of all transects and all critter sabers so we can calculate contacts between them

```{r}
list_of_matches <- array(NA,dim = c(2,2,number_of_animal_sabers_in_a_park))
list_of_surface_lines <- array(NA,dim = c(2,2,6))

for(i in 1:number_of_animal_sabers_in_a_park){
  list_of_matches[,,i] <- critter_saber()
}

line_placement <- seq(0,100, by=20)
for(i in 1:6){
  list_of_surface_lines[,,i] <- surface_line(line_placement[i])
}
```



```{r, echo=FALSE, fig.cap="After the angle is set, translate the lightsaber to a random location on the battlefield."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-1,101), ylim=c(-1,101))

park <- readPNG("park.png")
lim <- par()
rasterImage(park, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])

for(i in 1:6){
lines(list_of_surface_lines[,,i], col=adjustcolor("white", 1), lwd=3)

}

for(i in 1:number_of_animal_sabers_in_a_park){
lines(list_of_matches[,,i], col=adjustcolor("#00bcff", 0.2), lwd=2)
}

```

Calculate which sabers crossed my transect.
```{r}
crosses <- matrix(NA, 1, 10, byrow=TRUE)

for(i in 1:dim(list_of_matches)[3]){
for(j in 1:dim(list_of_surface_lines)[3]){
  crosses <- rbind(crosses, c(
    line.line.intersection(list_of_matches[1,,i], list_of_matches[2,,i], 
                           list_of_surface_lines[1,,j],     
    list_of_surface_lines[2,,j], interior.only = TRUE), 
    
    list_of_matches[1,,i], list_of_matches[2,,i], list_of_surface_lines[1,,j], 
    list_of_surface_lines[2,,j])
    )
  }
}

crosses <- crosses[which(is.na(crosses[,1]) == FALSE),]
colnames(crosses) <- c("x.cross", "y.cross", "x.p1", "y.p1", "x.p2", 
                    "y.p2", "x.surface.p1", "y.surface.p1", "x.surface.p2", 
                    "y.surface.p2")
head(crosses)
```


```{r, echo=FALSE, fig.cap="All the lightsabers we struck on my transect."}
par(mar=c(0,0,0,0))
plot(0,0,type="n", xaxt="n", yaxt="n", bty="n", xlab="", ylab="", 
     xlim=c(-1,101), ylim=c(-1,101))
rasterImage(park, lim$usr[1], lim$usr[3], lim$usr[2]-0.5, lim$usr[4])


for(i in 1:number_of_animal_sabers_in_a_park){
lines(list_of_matches[,,i], col=adjustcolor("white", 0.6), lwd=2)
}

for(i in 1:length(crosses[,1])){
lines(x = c(crosses[i,3], crosses[i,5]) , y = c(crosses[i,4], crosses[i,6]),  
      col=adjustcolor("#00bcff", 0.9), cex=1, pch=19, lwd=2)
}

for(i in 1:6){
lines(list_of_surface_lines[,,i], col=adjustcolor("white", 0.3), lwd=3)
}

```


Calculate Pi

### We take the ratio of the lightsabers we made contact with againt the ones we missed and we get a great approximation of Pi! 

```{r}
pi_estimate <- (2* light_saber_length * number_of_animal_sabers_in_a_park)/(distance_between_transects * length(crosses[,1]))
options(digits=20)
pi_estimate
```

# Happy Pi Day!